---
title: "Rmarkdown activity"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)
library(downloadthis)
```

## Rmarkdown activity  
![](https://youtu.be/PvfFr-eY77E){width="90%"}  
  
Data science projects in R can use a special data file type called an Rmarkdown (Rmd) file. To learn more about this file type and how to use it in Rstudio,  
  
1. Create a new project in Rstudio  
2. Download the rmarkdown activity file (click below)  
3. Move it to an Rstudio project directory  
4. Open the file and follow the instructions
5. If you get stuck at anytime, unpause the video above for a walkthrough

```{r, echo = FALSE}
#make temp file
temp_rmd <- tempfile(pattern = "1-rmarkdown_activity_file_", 
                     fileext = ".Rmd")
#download file to temp file
download.file(url = "https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/main/code/1-objects/1-rmarkdown_activity_file.Rmd", 
              destfile = temp_rmd)

#download button
downloadthis::download_file(
  path = temp_rmd,
  button_label = "Download Rmd",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

