---
title: "Introduction Final Exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)

# put back down below
# ![](https://youtu.be/XXXXXX){width="90%"}  
#   
# After watching the video, download the exercise Rmd file and move it to your project directory. Open it and follow the instructions. In this video, the exercise tasks will be described for you to carry out.
```

## Optional Exercises

Exercises are opportunities to practices several concepts from a module. Most will take 15-20 minutes to complete. These exercises below become progressively more challenging. Choose the one that suits your experience level, or work through them in order. Click the download button and then drag the exercise file to your working project directory.   
  
```{r}
h4("Cars Example")
make_button(file_name = "1-intro_exercise_file_mtcars.Rmd", 
            button_style = "primary")
br()
br()
h4("Heart Example")
make_button(file_name = "1-intro_exercise_file_heart.Rmd", 
            button_style = "primary")
br()
br()
h4("Subcell Example")
make_button(file_name = "1-intro_exercise_file_subcell.Rmd", 
            button_style = "primary")
actionLink("class1_exercise", "", icon = icon("bug"))

conditionalPanel(
  condition = "input.class1_exercise != 0",
  hr(),
  p("If you are participating in tidybiology live, then this is the file we will use in class. If you are not in a synchronous class, then feel free to use this additional file for practice."), 
  tags$ol(
    tags$li("Create a new project in Rstudio"), 
    tags$li("Download the project Rmd file (big red button)"), 
    tags$li("Follow along with the exercise")
  ),
  h4("In-class Example"),
  make_button(file_name = "1-intro_exercise_file_chromosome.Rmd",
              button_style = "danger"), 
  hr()
)
```




