---
title: "Exercise: Heart"
author: "Your Name Here"
date: "today"
---
## Load libraries
```{r, setup}
library(tidyverse)
```

## Import data
This is the code chunk that imports data. We will discuss how this code works in the next few sessions. All you need to know for now is that this imports an `R object` into your environment. Look to the right at the "Environment" tab and see a new object called `heart`. 
```{r import}
heart_url <- "https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/main/data/heart.csv"
heart <- read_csv(heart_url)
```

## Take a `glimpse` into the kinds of data in the dataframe
The `heart` object is a table (also called a dataframe), that has a series of columns and rows. The function `glimpse()` perfroms a quick peek at the data. We can see that this table has 15 columns (called 'Variables'), and 1025 rows (called 'Observations'). The name of each Variable is listed below (e.g. 'id', etc.). Importantly, R cares about capitalization, so `age` is NOT the same as `AGE`. The type of variable is listed next (e.g. <fct>, <dbl>, etc.), as we discussed. Lastly, the first few observations are listed.  
  
```{r glimpse}
glimpse(heart)
```

Another way to View the `heart` table object is to run the `View()` function. This is one of the odd functions in R that starts with a capital letter. Viewing the table will open a new tab in Rstudio to see all the data. This works great for small tables (like this one), but will crash R with big tables (100,000 x 100,000). To view this table, type in the console below: View(heart)  
  
Understanding where your data come from, and how your data are structured is one of the most important things in Data Science. This table was downloaded from: https://www.kaggle.com/johnsmith88/heart-disease-dataset/version/2. Go to a web browser to look at the original source.  
  
## My first graph  
Now we will make your first graph. Run the code chunk below to generate the image, and then continue reading about what each part of the code does.  
`ggplot` is the function  
`geom_point` is the type of graph - scatterplot with points, in this case  
`labs` is the layer for **lab**els  
`theme_minimal` is an easy layer to change several of the visual features of the graph  
`NULL` doesn't add anything to the graph, but allows you to comment out (#) layers without breaking the graph; if you break the graph and R is not responding, it's likely that `R` is waiting for another command; just hit `esc` on your keyboard to start over.  

```{r first_graph}
ggplot(heart, aes(x = age, y = chol, color = as.factor(sex))) +
  geom_point(size = 3) +
  labs(x = "Age", y = "Cholesterol", color = "Sex", title = "Cholesterol values from the heart disease dataset", caption = "Data from Kaggle | Plot from @matthewhirschey") +
  scale_colour_manual(labels = c("sex0", "sex1"), values = c("orange", "green")) +
  theme_minimal() +
  NULL
```

## This exercise has two final parts:  
1. One-by-one, comment out lines in the code chunk above (remember that # placed before a line of code comments it out). Figure out what each line of code does.
  
```{r quiz}
#Q: Which line is responsible for giving the orange/green color on the graph?  
#A: 
```

2. In this data set, the sex of the patient is coded as either 0 or 1. Go to the original data source (https://www.kaggle.com/johnsmith88/heart-disease-dataset/version/2) to determine which is male, and which is female. 
  
```{r quiz2}
#Q: What number is used for female?  
#A: 
```
  
3. Now that you know which is male and which is female, let's update the graph to make it more informative. First, copy the code that makes the plot and put it in the code chunk below. Click the green arrow to run it to make sure you can make the same graph again. Then, update the labels "sex0" and "sex1" to reflect male or female. Run the code chunk again.  
  
Finally, let's change the colors to pink and blue for female and male, respectively. We do this for two reasons. First, orange and green are not color-blind friendly, and many types of colorblindness prevents people from distinguishing between these colors. Second, in many cultures, pink and blue are intuitively connected to male and female. The more often you can make your visualization intuitive, the better.
  
```{r better_graph}
#copied code goes here

```
  
4. Make your own graph in the code chunk below. Use the code from the `{r first_graph}` code chunk. Copying and pasting is OK! Look at the different variables that are available for plotting by inspecting the `{r glimpse}` code chunk. The data dictionary for these variables is located at: https://www.kaggle.com/johnsmith88/heart-disease-dataset/version/2.  
  
Feel free to experiment and plot different information.  
- change the variable, if you'd like; the original plot compared heart id # vs. Basepairs, but 12 other variables are unused.  
- change the color of the points (be resourceful to find color names useable in `R`)   
- change the size of the points to something bigger
- finalize the labels (title, axes, caption) based on what you changed, and to reflect that you made it!  
* get creative; use cheatsheets or other resources to change other elements of this graph  
 
```{r best_graph}
#your code goes here

```