---
title: "Exercise: mtcars"
---
## Load libraries
```{r setup}
# most packages are on a central repository called CRAN, which makes them easy to install. If you have not yet installed the tidyverse, uncomment the line below (by removing the # at the beginning) to get the tidyverse packages.  You will also get a prompt to install a few more packages that are required for the tidyverse. Just type "Yes" when asked. If it successfully installed, then you can run the last line of this code chunk to load the library.

# install.packages("tidyverse")
library(tidyverse)
```

## Import data
This is the code chunk that imports data. We will discuss how this code works in the next few sessions. All you need to know for now is that this imports an `R object` into your environment. `mtcars` is a dataset that is included with R. Look to the right at the "Environment" tab and see a new object called `mtcars`. 
```{r import}
data("mtcars")
```

## Take a `glimpse` into the kinds of data in the dataframe
The `mtcars` object is a table (also called a dataframe), that has a series of columns and rows. The function `glimpse()` performs a quick peek at the data. We can see that this table has 11 columns (called 'Variables'), and 32 rows (called 'Observations'). The name of each Variable is listed below (e.g. 'mpg', 'cyl', etc.). Importantly, R cares about capitalization, so `mpg` is NOT the same as `MPG`. The type of variable is listed next (e.g. <fct>, <dbl>, etc.), but you don't need to worry about these yet. Lastly, the first few observations are listed.  
  
```{r glimpse}
glimpse(mtcars)
```

Another way to View the `mtcars` table object is to run the view function. This is one of the odd functions in R that starts with a capital letter. Viewing the table will open a new tab in Rstudio to see all the data. This works great for small tables (32x11), but will crash R with big tables (100,000 x 100,000). To view this table, type in the console below: View(mtcars)  
  
Understanding where your data come from, and how your data are structured is one of the most important things in Data Science. The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973--74 models).  
  
## My first graph  
Now we will make your first graph. Run the code chunk below to generate the image, and then continue reading about what each part of the code does.  
`ggplot` is the function  
`geom_point` is the type of graph - scatterplot with points, in this case  
`labs` is the layer for **lab**els  
`theme_minimal` is an easy layer to change several of the visual features of the graph  
`scale_y_continuous` allows me to change the scale to make it more readable  
`NULL` doesn't add anything to the graph, but allows you to comment out (#) layers without breaking the graph; if you break the graph and R is not responding, it's likely that `R` is waiting for another command; just hit `esc` on your keyboard to start over.  

```{r first_graph}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point(color = "darkblue", size = 3) +
  geom_smooth(method='loess') +
  labs(x = "Displacement (cu.in.)", y = "Miles/(US) gallon", title = "Bigger Engines Are Less Efficient", caption = "Data from mtcars | Plot from @hirscheylab") +
  theme_minimal() +
  NULL
```

## This project has two final parts:  
1. One-by-one, comment out lines in the code chunk above. Figure out what each line of code does.
  
```{r quiz}
#Q: Which line is responsible for adding a trendline on the graph?  
#A: 
```

2. Make your own graph in the code chunk below. Use the code from the `{r first_graph}` code chunk. Copying and pasting is OK! Look at the different variables that are available for plotting by inspecting the `{r glimpse}` code chunk. Feel free to experiment and plot different information.  
* change the variable, if you'd like; the original plot compared mtcars disp vs. mpg, but 9 other variables are unused.  
* change the color of the points (be resourceful to find color names useable in `R`)   
* change the size of the points to something bigger
* finalize the labels (title, axes, caption) based on what you changed, and to reflect that you made it!  
* get creative; use cheatsheets or other resources to change other elements of this graph  
 
```{r second_graph}
#your code goes here

```