---
title: "Filtering Observations"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
library(gradethis)
gradethis::gradethis_setup()

happy_full <- 
  read_csv("https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/main/data/world-happiness-report-2021.csv") %>% 
  clean_names() %>% 
  rename(region = regional_indicator)
```
## Filtering exercises

`happy_full` contains both numeric and character variables, with lots and lots of observations (rows). This gives us a great opportunity to practice our filtering skills!

### Simple filters

Say we're only interested in looking at data for countries in East Asia. How would we do this?

```{r simple-filter, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Simple filter"}

```

<div id="simple-filter-hint">
**Hint:** To see which major regions are in this dataset, run `unique(happy_select$region)`
</div>

```{r simple-filter-solution}
happy_select %>% 
  filter(region == "East Asia")
```

```{r simple-filter-code-check}
grade_code()
```

### Filter%in%g

Now use `filter()` to only keep data for the following countries - Algeria, Belgium, India, Tunisia, and Uganda. Try to do this without writing multiple filter statements  

```{r multiple-filter, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="multiple-filter-hint">
**Hint:** Try the `%in%` operator
</div>

```{r multiple-filter-solution}
happy_select %>% 
  filter(country_name %in% c("Algeria", "Belgium", "India", "Tunisia", "Uganda"))
```

```{r multiple-filter-code-check}
grade_code()
```

### Numeric filtering

Finally, let's filter out information for countries that have a **below average** ladder score

```{r num-filter, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="num-filter-hint">
**Hint:** Use the base R `mean()` function
</div>

```{r num-filter-solution}
happy_select %>% 
  filter(ladder_score > mean(ladder_score))
```

```{r num-filter-code-check}
grade_code()
```