---
title: "Selecting Variables Exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
library(gradethis)
gradethis::gradethis_setup()

happy_full <- 
  read_csv("https://raw.githubusercontent.com/BAREJAA/website_for_john/master/datasets/world-happiness-report-2021.csv") %>% 
  clean_names() %>% 
  rename(region = regional_indicator)
```
## Select exercises

We have preloaded a dataset for you called `happy_full`. This dataset contains many variables. This gives us the chance to practice our `select()`-ing skills!

### Simple selects

Let's warm up by performing some basic select operations. The coding console below is the same type as you'll find in Rstudio. Because the data are already loaded, you can explore it.    

How would you select just the columns `region` and `ladder_score`?  

```{r simple-select, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "A simple selection"}

```

```{r simple-select-solution}
happy_full %>% 
  select(region, ladder_score)
```

```{r simple-select-code-check}
grade_code()
```

Now select everything between (and including) `social_support` and `generosity`

```{r select-between, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Selecting multiple columns"}

```

```{r select-between-solution}
happy_full %>% 
  select(social_support:generosity)
```

```{r select-between-code-check}
grade_code()
```

### Slightly-more-difficult selects  

Let's try something more challenging now. Select all variables that **do not** have underscores in their names

```{r no-underscore, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "No underscores"}

```

<div id="no-underscore-hint">
**Hint:** You'll need a helper function. Also, don't forget `!`
</div>

```{r no-underscore-solution}
happy_full %>% 
  select(!contains("_"))
```

```{r no-underscore-code-check}
grade_code()
```