---
title: "dplyr Exercise: proteins"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Loading the `proteins` dataset into RStudio 

The dataset can be found in the `proteins` package on the Hirscheylab Github page

To install and load the packages, uncomment the install code, run the install command, and then re-comment out. Make sure you have `devtools` installed, and if not, install that one first from CRAN.  
```{r install, echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE}
# devtools::install_github("hirscheylab/proteins")
library(proteins)
```

## Inspecting the dataset
Use the `dim()` function to see how many rows (observations) and columns (variables) there are
```{r dim, echo = TRUE, warning = FALSE}

```

## Inspecting the dataset
Use the `glimpse()` function to see what kinds of variables the dataset contains
```{r glimpse, echo = TRUE, warning = FALSE}

```

## Basic Data Types in R 
R has 6 basic data types -

**character** - `"a"`, `"tidyverse"`  

**numeric** - `2`, `11.5`  

**integer** - `2L` (the `L` tells R to store this as an integer)  

**logical** - `TRUE`, `FALSE`  

**complex** - `1+4i`  

(**raw**)

You will also come across the **double** datatype. It is the same as **numeric**

**factor**. A **factor** is a collection of *ordered* character variables

In addition to the `glimpse()` function, you can use the `class()` function to determine the data type of an object or of part of an object (like a specific column). Try calling `class` on proteins or one of its vectors.
```{r class, echo = TRUE, warning = FALSE}

```

## (Re)Introducing `%>%` 
The `%>%` operator is a way of "chaining" together strings of commands that make reading your code easy. To practice how the pipe works, start with the proteins dataframe, "and then" select `uniprot_id` and `length`, "and then" filter `length` greater than 500
```{r pipe, eval = TRUE, echo = TRUE}

```

What would the nested alternative to using `%>%` be? 
```{r alternative, eval = FALSE, echo = TRUE}

```

Although this is only one line as opposed to three, it's both more difficult to write and more difficult to read

## Introducing the main dplyr verbs
dplyr is a package that contains a suite of functions that allow you to easily manipulate a dataset. Some of the things you can do are:  
- select rows and columns that match specific criteria  
- create new variables (columns)  
- obtain summary statistics on individual groups within your datsets  
  
The main verbs we will review in this exercise are `select()`, `filter()`, `arrange()`, `mutate()`, and `summarise()`. These all combine naturally with `group_by()` which allows you to perform any operation "by group"

## `select()` 
The `select()` verb allows you to extract specific columns from your dataset. The most basic `select()` is one where you comma separate a list of columns you want included. For example, try to select the `uniprot_id` and `length` columns
```{r select_simple, eval = TRUE, echo = TRUE}

```

If you want to select all columns *except* `uniprot_id`, how would you do that?
```{r select_exclude, eval = TRUE, echo = TRUE}

```

Finally, you can provide a range of columns to return two columns and everything in between. For example, try using the proteins dataframe and selecting `uniprot_id`, `protein_name`, and everything in-between
```{r select_range, eval = TRUE, echo = TRUE}
 
```

From proteins, select the following columns - `uniprot_id`, `sequence`, `length`, and `mass`
```{r select_solution, eval = TRUE, echo = TRUE}

```

## `filter()` 
The `filter()` verb allows you to choose rows based on certain condition(s) and discard everything else. All filters are performed on some logical statement. If a row meets the condition of this statement (i.e. is true) then it gets chosen (or "filtered"). All other rows are discarded. Remember, `filter()` only applies to rows, and has no effect on columns.

First, filtering can be performed on categorical data, such as a single gene within the `proteins` dataframe. Try filtering for the gene "TP53".
```{r filter_categorical, eval = TRUE, echo = TRUE}

```

Filtering can also be performed on numerical data. For example, try to identify proteins with a `length` that is shorter than 50 (small proteins!)
```{r filter_numerical, eval = TRUE, echo = TRUE}
 
```

To filter on multiple conditions, you can write a sequence of `filter()` commands. For example, to choose proteins with a `protein_name` "Collagen alpha-1" **and** a `length` longer than 1000. 
```{r filter_both, eval = TRUE, echo = TRUE}

```

To avoid writing multiple `filter()` commands, multiple logical statements can be put inside a single `filter()` command, separated by commas. Try rewriting the same code from the chunk above and placing it here.
```{r filter_both_efficient, eval = TRUE, echo = TRUE}
 
```

Filter all proteins with a protein name that is **not** equal to "Collagen alpha-1" **and** a `length` that is **greater** than or equal to 1000

`!=` = "not equal to"

`>=` = "less than or equal to"
```{r filter_exercise, eval = TRUE, echo = TRUE}
 
```

## `arrange()` 
You can use the `arrange()` verb to sort rows. The input for arrange is one or many columns, and `arrange()` sorts the rows in ascending order i.e. from smallest to largest. Sort rows from smallest to largest protein length.
```{r arrange, eval = TRUE, echo = TRUE}

```

To reverse this order, use the `desc()` function within `arrange()`. Try arranging the protein lengths from largest to smallest now.
```{r arrange_desc, eval = TRUE, echo = TRUE}
 
```

What happens when you apply `arrange()` to a categorical variable? Try arranging on one of the categorical variables withing `proteins`
```{r arrange_exercise, eval = TRUE, echo = TRUE}

```

## `mutate()` 
The `mutate()` verb, unlike the ones covered so far, creates new variable(s) i.e. new column(s). For example, try taking all the elements of the column `length`, evaluate the square root of each element, and populates a new column called `sqrt_length` with these results
```{r mutate_simple, eval = TRUE, echo = TRUE}
 
```

Multiple columns can be used as inputs. For example, try taking the `length` and divide my `mass` and storing the results in a new column called `protein_length_mass`
```{r mutate_multiple, eval = TRUE, echo = TRUE}

```

Create a new column (give it any name you like) and fill it with protein lengths divided by 100
```{r mutate_exercise, eval = TRUE, echo = TRUE}
 
```

## `summarise()` 
`summarise()` produces a new dataframe that aggregates that values of a column based on a certain condition. For example,  calculate the mean protein length and mass
```{r summarise, eval = TRUE, echo = TRUE}

```

You can assign your own names by using the same code organization we saw in `mutate()`. Try taking the same code you wrote above to summarize the mean(length) and mean(mass), but set these results to equal `mean_length` and `mean_mass`, respectively.
```{r summarise_name, eval = TRUE, echo = TRUE}

```

Now make a new table that contains the mean, median and standard deviations of protein lengths. Use the `median()` and `sd()` functions to calculate median and standard deviation
```{r summarise_exercise, eval = TRUE, echo = TRUE}

```

## `group_by()` 
Let's practice `group_by()` and `summarise()`. These functions are most often used in combination to summarise by groups. For example, if you'd like to know the average length based on the first letter of their gene name (silly example), you would first create a new column with that letter, then `group_by()` the `first_letter` variable, and finally then calculate the mean length. The code chunk is pre-populated withe the code to extract the first letter of each gene name to help.
```{r group_by_and_summarise, eval = TRUE, echo = TRUE}
proteins %>% 
  mutate(first_letter = stringr::str_extract(gene_name, pattern = "[[:alpha:]]"))
```

## Saving a new dataset 
If you'd like to save the output of your wrangling, you will need to use the `<-` or `=` operators. Try taking the code you wrote in the last code chunk and now modify it here to create a new object called `size_by_letter`
```{r save, eval = TRUE, echo = TRUE}
 
```

Finally, save `size_by_letter` as a new file (e.g. csv)
```{r csv_save, eval = FALSE, echo = TRUE}

```

## For continued learning
Run the following to access the `dplyr` vignette for more examples
```{r Dplyr vignette, eval = FALSE, echo = TRUE}
browseVignettes("dplyr")
```



