---
title: "Programmatic Access"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

Is there a way to programmatically access the help pages for functions?
Why do this? To build up a question bank for the quiz app

From - https://community.rstudio.com/t/programmatically-accessing-param-fields-values-of-functions-from-installed-package/14731

Q - "is there a way to assign to a global environment variable the @param field value of a function in an installed package, via ({tools} or {utils})?"

```{r}
package_docs <- function(pkg) {
  help_dir <- system.file("help", package = pkg)
  db_path <- file.path(help_dir, pkg)
  tools:::fetchRdDB(db_path)
}

# e.g.
rdb <- package_docs("ggplot2")
# rdb is a large list
names(rdb) # 193 entries, corresponding to 193 functions

rdb[[2]][[2]][[1]][1] # this is how you get the function name
rdb[[2]][[1]][[1]][1] # this is how you get the function description

# url
# https://www.rdocumentation.org/packages/ggplot2/versions/3.3.6/topics/add_theme

# to get package version number - 
packageVersion("ggplot2")
```

Detecting parts of speech

Goal - only select functions whose descriptions begin with a verb so that you can have questions of the form - 
"How would you..."
```{r}
library(udpipe)
ud_model <- udpipe_download_model(language = "english")
ud_model <- udpipe_load_model(ud_model$file_model)
x <- as.data.frame(udpipe_annotate(ud_model, x = "summarise"))
x # this doesn't work so well...

word_list <- c()
for(i in 1:length(rdb)){
  word_list[i] <- word(rdb[[i]][[1]][[1]][1])
}

unique(word_list)
```

