---
title: "First r exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)
```

## R Exercise
### Activity 1
This code computes the answer to one plus one.  
1. Run the code by clicking the button and watch for the answer to appear below.  
2. Next, change it so it computes two plus two, and check your answer:  

```{r addition, exercise=TRUE}
1 + 1
```
  
### Activity 2
This box is called a Console and you can run R code within it. Try a few other simple math operations.

```{r addition2, exercise=TRUE}

```
  
### Activity 3
Console are powerful and can run any R code within it. We've populated the console below with some code. Run the code to see what it does.

```{r advanced_example, exercise=TRUE}
#view the first 5 rows of the 'mtcars' dataset
head(mtcars, n = 5)

#build a plot from the 'mtcars' dataset
ggplot(mtcars, aes(x = mpg, y = hp, color = as_factor(cyl))) + 
geom_point() +
labs(color = "Cylinders") +
theme_minimal()
```

