---
title: "Rmarkdown activity"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)
```

## Rmarkdown activity
Data science projects in R can use a special data file type called an Rmarkdwon (Rmd) file. To learn more about this file type and how to use it in Rstudio,  
  
1. Download the [rmarkdown activity file](www/objects/6-rmarkdown_activity_file.Rmd)  
2. Move it to an Rstudio project directory  
3. Open the file and follow the instructions  
-- Inspect the code starting with the set-up chunk. Below it, you’ll see code that imports data, stores objects, and creates a graph.  
-- Knit the document and see the results; pretty messy, huh?  
-- Use what you've learn to execute the code, but not include the results of the set-up chunk and data import chunk.  
-- Replace every **BOLD** with inline R code  
-- Remove the {calculations} chunk so it is not included with the output  
-- Change the {graph} chunk so that only the output of the plot is shown, but not the code.  
-- Re-Knit the document  
-- **Beautiful!**  
4. If you get stuck, or want to check your answers, download the answer file below
  
```{r, echo = FALSE}
HTML("<a href='www/objects/6-rmarkdown_activity_file_answers.Rmd'>Check your answers</a>")
```

