---
title: "Parameterization exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)
library(downloadthis)
```

## Parameter exercise
Parameters of a document are defined in the YAML header, and can pre-populate an Rmarkdown document. To see this in action, download this file for a parameterization activity.
  
1. Download the parameter activity file below    
2. Move it to an Rstudio project directory  
3. Open the file, and follow these instructions  
-- Click the dropdown menu next to Knit and use Knit with Parameters to render the document  
-- Change the parameter values to match your age and gender, and re-knit.
  
```{r, echo = FALSE}
#make temp file
temp_rmd <- tempfile(pattern = "6-parameter_activity_file_", 
                     fileext = ".Rmd")
#download file to temp file
download.file(url = "https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/master/code/6-objects/6-parameter_activity_file.Rmd", 
              destfile = temp_rmd)

#download button
downloadthis::download_file(
  path = temp_rmd,
  button_label = "Download Rmd",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```

