---
title: "rmarkdown Final Exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)

```

## Optional Exercises

Exercises are opportunities to practices several concepts from a module. If you did not follow along during the videos, then download these files and complete them on your own. These should take ~30 minutes to finish. Work through the exercises, downloading the answers if needed.  
  
To begin, click the download button and then drag the exercise file to your working project directory. Again, the "answer" file shows you one way that you can accomplish the task.   
  
```{r}
h4("rmarkdown Activity")
make_button(file_name = "6-rmarkdown_activity_file.Rmd", 
            button_style = "primary")
make_button(file_name = "6-rmarkdown_activity_file_answers.Rmd",
            button_style = "default")
br()
br()
h4("Parameter Activity")
make_button(file_name = "6-parameter_activity_file.Rmd", 
            button_style = "primary")
make_button(file_name = "6-parameter_activity_file_answers.Rmd",
            button_style = "default")
actionLink("class6_exercise", "", icon = icon("bug"))

conditionalPanel(
  condition = "input.class6_exercise != 0",
  hr(),
  p("If you are participating in tidybiology live, then this is the file we will use in class. If you are not in a synchronous class, then feel free to use this additional file for practice."), 
  tags$ol(
    tags$li("Create a new project in Rstudio (optional)"), 
    tags$li("Download the project Rmd file (big red button)"), 
    tags$li("Follow along with the exercise")
  ),
  h4("In-class Example"),
  make_button(file_name = "6-rmarkdown_exercise_file_cmv.Rmd",
              button_style = "danger"), 
  make_button(file_name = "6-rmarkdown_exercise_file_cmv_answers.Rmd",
              button_style = "default"), 
  hr()
)
```
