---
title: "Rmarkdown activity"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)
library(downloadthis)
```

## Rmarkdown activity
Data science projects in R can use a special data file type called an Rmarkdwon (Rmd) file. To learn more about this file type and how to use it in Rstudio,  
  
1. Download the rmarkdown activity file below  
2. Move it to an Rstudio project directory  
3. Open the file and follow the instructions  
-- Inspect the code starting with the set-up chunk. Below it, you’ll see code that imports data, stores objects, and creates a graph.  
-- Knit the document and see the results; pretty messy, huh?  
-- Use what you've learn to execute the code, but not include the results of the set-up chunk and data import chunk.  
-- Replace every **BOLD** with inline R code  
-- Remove the {calculations} chunk so it is not included with the output  
-- Change the {graph} chunk so that only the output of the plot is shown, but not the code.  
-- Re-Knit the document  
-- **Beautiful!**  
4. If you get stuck, or want to check your answers, download the answer file below
  
```{r, echo = FALSE}
#make temp file
temp_rmd <- tempfile(pattern = "6-rmarkdown_activity_file_", 
                     fileext = ".Rmd")
#download file to temp file
download.file(url = "https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/master/code/6-objects/6-rmarkdown_activity_file.Rmd", 
              destfile = temp_rmd)

#download button
downloadthis::download_file(
  path = temp_rmd,
  button_label = "Download Rmd",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)

#make temp file
temp_rmd_answers <- tempfile(pattern = "6-rmarkdown_activity_file_answers_", 
                     fileext = ".Rmd")
#download file to temp file
download.file(url = "https://raw.githubusercontent.com/matthewhirschey/tidybiology.org/master/code/6-objects/6-rmarkdown_activity_file_answers.Rmd", 
              destfile = temp_rmd_answers)

#download button
downloadthis::download_file(
  path = temp_rmd_answers,
  button_label = "Download Answers",
  button_type = "primary",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = TRUE
)
```
  