---
title: "Advanced Filter"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
gradethis::gradethis_setup()
```
## Advanced Filter Exercise

We will use the `happy_full` data set for these exercises

Choose countries that are NOT in Europe i.e. countries that have a `region` entry that is NOT equal to either "Western Europe" or "Central and Eastern Europe"
```{r filter-in, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Boxplot"}

```

<div id="filter-in-hint">
You can use the `%in%` symbol
</div>

```{r filter-in-solution}
happy_full %>% 
  filter(!region %in% c("Western Europe", "Central and Eastern Europe")) 
```

```{r filter-in-code-check}
grade_code()
```

Choose countries that do NOT have an above average `healthy_life_expectancy` or above average `generosity` score

```{r filter-or, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="filter-or-hint">
Use the `|` operator
</div>

```{r filter-or-solution}
happy_full %>% 
  filter(!(healthy_life_expectancy > mean(healthy_life_expectancy)|
         generosity > mean(generosity))) 
```

```{r filter-or-code-check}
grade_code()
```
