---
title: "Between and Near"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
gradethis::gradethis_setup()
```
## Between and Near Exercise

We will again use the `happy_full` data set for these exercises

Choose countries that have a `generosity` score that is between the median and mean `generosity` scores (inclusive) across all countries
```{r filter-between, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Boxplot"}

```

```{r filter-between-solution}
happy_full %>% 
  filter(between(generosity, median(generosity), mean(generosity)))
```

```{r filter-between-code-check}
grade_code()
```

Choose countries whose `ladder_score` is equal to the mean `ladder_score` +/- the standard error of `ladder_score` across all countries

```{r filter-near, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="filter-near">
Standard error formula - sd/sqrt(n)
</div>

```{r filter-near-solution}
happy_full %>% 
  filter(near(ladder_score, mean(ladder_score), tol = sd(ladder_score)/sqrt(rows))) 
```

```{r filter-near-code-check}
grade_code()
```
