  
### R Markdown
Plain text file with 3 types of content  
```{r rmarkdown_3parts, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown.png"
knitr::include_graphics(url)
```

### Use the 'knit' button to render a report
```{r rmarkdown_knit_button, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/96ec0c54c6d64ea2ec3665db9b3b781962ff6339/5cee1/lesson-images/how-3-output.png"
knitr::include_graphics(url)
```

### How it works
```{r rmarkdown_knit, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://d33wubrfki0l68.cloudfront.net/61d189fd9cdf955058415d3e1b28dd60e1bd7c9b/b739c/lesson-images/rmarkdownflow.png"
knitr::include_graphics(url)
```
  
When you run render, R Markdown feeds the .Rmd file to knitr ⧉, which executes all of the code chunks and creates a new markdown (.md) document which includes the code and it’s output.  
  
The markdown file generated by `knitr` is then processed by `pandoc` which is responsible for creating the finished format.  
  
This may sound complicated, but R Markdown makes it extremely simple by encapsulating all of the above processing into a single render function.  
  
### Key `knitr` points {.build}
- `Knitr` runs the document in a fresh R session, which means you need to load the libraries that the document uses **in the document**  
  
- Objects made in one code chunk will be available to code in later code chunks, but not before  
  
- For example, first create `dataframe_name` and then using `dplyr::left_join` you create `dataframe_joined_name`, `dataframe_name` will be available later on in the document to do this. **However, you cannot use `dataframe_joined_name` in a code chunk before you make it, even if it is available in your environment**

- To keep this straight, just think (and code) in sequential chunks

### Rmarkdown -- The definitive Guide
[https://bookdown.org/yihui/rmarkdown/](https://bookdown.org/yihui/rmarkdown/)  
<center><iframe src="https://bookdown.org/yihui/rmarkdown/"></iframe></center>


### Markdown is a simplifed language to format text
```{r rmarkdown_text, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://d33wubrfki0l68.cloudfront.net/59f29676ef5e4d74685e14f801bbc10c2dbd3cef/c0688/lesson-images/markdown-1-markup.png"
knitr::include_graphics(url)
```

### Markdown can 'markup' text to be {.build}
- **bold**  
- _italics_  
- `code`  
- unformatted text  
- ### Medium headers (3)  
- ###### Small headers (6)  
- Bullets (like this list)  
- [Links, such as to the github repository storing class material](https://github.com/matthewhirschey/)  
- ...and many more
  
### Rmarkdown cheatsheet & quick reference
<center><iframe src="https://rmarkdown.rstudio.com/lesson-15.html"></iframe></center>
  
### Rmarkdown <br>Code
Any questions?  

### Rmarkdown Code
```{r rmarkdown_code, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Shortcut
```{r rmarkdown_code_shortcut, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-code-shortcut.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk1, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk1.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk2, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk2.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk3, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk3.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk4, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk4.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk5, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk5.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk6, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk6.png"
knitr::include_graphics(url)
```

### Rmarkdown Code Chunk Options
```{r rmarkdown_chunk7, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-chunk7.png"
knitr::include_graphics(url)
```

### Clean-up the Rmarkdown exercise
1. Go to code/  
2. Open 10_rmarkdown.Rmd  
3. Inspect the code starting with the set-up chunk. Below it, you’ll see code that imports data, stores objects, and creates a graph.  
4. Knit the document and see the results; pretty messy, huh?  
5. Use what you've learn to execute the code, but not include the results of the set-up chunk and data import chunk.  
6. Replace every **BOLD** with inline R code  
7. Remove the {calculations} chunk so it is not included with the output  
8. Change the {graph} chunk so that only the output of the plot is shown, but not the code.
9. Re-Knit the document  
10. **Beautiful!**  

### Rmarkdown header is called 'YAML'
```{r rmarkdown_yaml, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-yaml.png"
knitr::include_graphics(url)
```

### Rmarkdown output is defined here
```{r rmarkdown_output, echo=FALSE, fig.cap="", out.width='66%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-output.png"
knitr::include_graphics(url)
```

### Rmarkdown formats
Recall that Rmarkdown documents can be rendered into several different output file types
<center><iframe src="https://rmarkdown.rstudio.com/formats.html"></iframe></center>

### Rmarkdown Parameters
Parameters of a document are defined in the YAML header, and can pre-populate an Rmarkdown document. To see this in action,  
  
1. Open 11_rmarkdown_params.Rmd.  
2. Click the dropdown menu next to Knit and use Knit with Parameters to render the document.  
3. What happens if you type in a different gender or different age?  

--- 

```{r rmarkdown_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-parameters.png"
knitr::include_graphics(url)
```

--- 

```{r rmarkdown_using_params, echo=FALSE, fig.cap="", out.width='75%'}
url <- "https://github.com/matthewhirschey/tidybiology-plusds/raw/master/media/Rmarkdown-using-parameters.png"
knitr::include_graphics(url)
```

### Take another look
<h1>Re-inspect 11_rmarkdown_params.Rmd.</h1>
<br>
<br>
Where were the parameters called in the code?
