---
title: "stringr Exercise: mitocarta"
output: html_document
---

## Load package
For this activity, you'll need the following packages: tidyverse, tidybiology, and mitocarta. The last two are on github, so you'll need to install from there. Hint: devtools::github_install()
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# devtools::github_install("matthewhirschey/tidybiology")
# devtools::github_install("matthewhirschey/mitocarta")
library(tidyverse)
library(tidybiology)
library(mitocarta)
```

## Required packages and datasets
Load the datasets called proteins and mitocarta, and then take a quick glimpse to remind yourself what they look like. 
```{r}
data(proteins)
data(mitocarta)

glimpse(proteins)
glimpse(mitocarta)
```

## Clean the datasets
From the proteins dataset, keep only the `gene_name` and `sequence` columns. From the mitocarta dataset, keep only the `symbol` and `mito_domain_score` columns. Also, the tidyr (part of the tidyverse) package has a function called drop_na(), which will remove NAs from an indicated column. Use this function to drop NAs from `mito_domain_score`. Finally, save these new objects as the same name as the original datasets, thereby over-writing them in your environment. 
```{r}
# keep just a few columns in `proteins` and `mitocarta`
proteins <- 
  proteins %>% 
  select(gene_name, sequence)

mitocarta <- 
  mitocarta %>% 
  select(symbol, mito_domain_score) %>% 
  drop_na(mito_domain_score)
```

## Sequence Lengths
Let's practice some of the stringr functions. How would you get the sequence lengths for all the entries in the `proteins` dataframe? Tip: first take another `glimpse()` at this dataframe to see which column would be relevant to this question
```{r}
glimpse(proteins)
stringr::str_length(proteins$sequence)
```

Next, let's count the number of lysines in each sequence. Instead of using str_length, which essentially counts everything, we want to use str_count. This is going to count only the things that match a pattern argument. 
```{r}
str_count(string = proteins$sequence, 
          pattern = "K")
str_count(proteins$sequence, "K") #this is equivalent, just shorter
```

For each sequence, how many times does an arginine follow a lysine?

```{r}
str_count(proteins$sequence, "KR")
```

How would you test if the following motif - "RCP.L" - was in a particular sequence? How many proteins contain this motif? What does the "." mean?
```{r}
str_detect(proteins$sequence, "RCP.L")
sum(str_detect(proteins$sequence, "RCP.L"))
# "." represents any character
```

How would you create a new dataframe that only contains proteins that contain the "RCP.L" motif?
```{r}
proteins %>% 
  dplyr::filter(stringr::str_detect(sequence, "RCP.L"))
```

Identify proteins that contain an ATM Kinase target site, which is either SQ or TQ. Create a new column called `atm` that contains a `TRUE` if a protein has this motif and a `FALSE` otherwise
```{r}
proteins %>% 
  dplyr::mutate(atm = str_detect(sequence, "[S|T]Q")) %>% 
  dplyr::relocate(atm) # moves the named column to the start of the data frame; not required, but useful
```

What are the mean protein lengths for proteins that contain an ATM site versus proteins that do not? Hint: use the code you wrote in the last code chunk and build upon it.
```{r}
proteins %>% 
  mutate(atm = str_detect(sequence, "[S|T]Q")) %>% 
  group_by(atm) %>% 
  summarise(mean_length = mean(length))
```

Which columns in `proteins` and `mitocarta` can we use to join on?
```{r}
glimpse(proteins)
glimpse(mitocarta)
# We see that `gene_name` from `proteins` and `symbol` from `mitocarta` contain the same information, and so we use this to join on
```

Combine `proteins` and `mitocarta`, ensuring that there are no missing values in the `mito_domain_score` column of the resulting dataframe (call this new dataframe `mito_proteins`)
```{r}
mito_proteins <- 
  proteins %>% 
  right_join(mitocarta, by = c("gene_name" = "symbol")) 

mito_proteins %>% 
  count(mito_domain_score) # there are now no missing values in this column
```

Combine `proteins` and `mitocarta` in such a way that the resulting dataframe contains no missing values. Do not save the output but do confirm that there are no missing values in the output
```{r}
proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  nrow() # 15631

proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  drop_na() %>% 
  nrow() # also 15631, confirming that there are no missing values in the output
```

Great job!