---
title: "Stringr Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `proteins`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)
data(proteins)
```

## Please answer the following questions

How would you get the sequence lengths for all the entries in the `proteins` dataframe? Hint: first take a `glimpse()` at this dataframe to see which column would be relevant to this question

```{r}

```

Count the number of lysines in each sequence

```{r}

```

For each sequence, how many times does an arginine follow a lysine?

```{r}

```

How would you test if the following motif - "RCP.L" - was in a particular sequence? How many proteins contain this motif? What does the "." mean?

```{r}

```

How would you create a new dataframe that only contains proteins that contain the "RCP.L" motif?

```{r}

```

Identify proteins that contain the E-box motif - "CANNTG". Create a new column called `e_box` that contains a `TRUE` if a sequence has this motif and a `FALSE` otherwise

```{r}

```

What are the mean protein lengths for proteins that contain an E-box versus proteins that do not?

```{r}

```

