---
title: "Stringr Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `proteins`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
remotes::install_github("matthewhirschey/mitocarta", force = TRUE)
library(tidybiology)
data(proteins)
data(mitocarta)

# keep just a few columns in `proteins` and `mitocarta`
proteins <- proteins %>% 
              select(gene_name, sequence)
mitocarta <- mitocarta %>% 
              select(symbol, mito_domain_score) %>% 
              drop_na(mito_domain_score)
```

## Please answer the following questions

How would you get the sequence lengths for all the entries in the `proteins` dataframe? Hint: first take a `glimpse()` at this dataframe to see which column would be relevant to this question

```{r}
glimpse(proteins)
str_count(proteins$sequence)
```

Count the number of lysines in each sequence

```{r}
str_count(proteins$sequence, "K")
```

For each sequence, how many times does an arginine follow a lysine?

```{r}
str_count(proteins$sequence, "KR")
```

How would you test if the following motif - "RCP.L" - was in a particular sequence? How many proteins contain this motif? What does the "." mean?

```{r}
str_detect(proteins$sequence, "RCP.L")
sum(str_detect(proteins$sequence, "RCP.L"))
# "." represents any character
```

How would you create a new dataframe that only contains proteins that contain the "RCP.L" motif?

```{r}
proteins %>% 
  filter(str_detect(sequence, "RCP.L"))
```

Identify proteins that contain the E-box motif - "CANNTG". Create a new column called `e_box` that contains a `TRUE` if a sequence has this motif and a `FALSE` otherwise

```{r}
proteins %>% 
  mutate(e_box = str_detect(sequence, "CA..TG")) %>% 
  relocate(e_box) # this moves the named column to the start of the data frame
```

What are the mean protein lengths for proteins that contain an E-box versus proteins that do not?

```{r}
proteins %>% 
  mutate(e_box = str_detect(sequence, "CA..TG")) %>% 
  group_by(e_box) %>% 
  summarise(mean_length = mean(length))
```

Which columns in `proteins` and `mitocarta` can we use to join on?

```{r}
glimpse(proteins)
glimpse(mitocarta)
# We see that `gene_name` from `proteins` and `symbol` from `mitocarta` contain the same information, and so we use this to join on
```

Combine `proteins` and `mitocarta`, ensuring that there are no missing values in the `mito_domain_score` column of the resulting dataframe (call this new dataframe `mito_proteins`)

```{r}
mito_proteins <- proteins %>% 
                  right_join(mitocarta, by = c("gene_name" = "symbol")) 
mito_proteins %>% 
  count(mito_domain_score) # there are now no missing values in this column
```

Combine `proteins` and `mitocarta` in such a way that the resulting dataframe contains no missing values. Do not save the output but do confirm that there are no missing values in the output

```{r}
proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  nrow() # 15631
proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  drop_na() %>% 
  nrow() # also 15631, confirming that there are no missing values in the output
```
