---
title: "ggplot Activity"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
gradethis::gradethis_setup()
```
## ggplot Activity

The `midwest` data set contains multiple variables. For these exercises will just be using `state` and `area`. If you want to `glimpse` at the data again, here is an extra scratchpad.

### Scratchpad
```{r ggplot_activity_scratchpad, exercise=TRUE}

```

### Activity 1: A boxplot

Make a boxplot to plot the distribution of land area for each state  

NOTE: Use the verbose form of ggplot code i.e. Remember to include `data = ` and `mapping = `

```{r boxplot, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Boxplot"}

```

<div id="boxplot-hint">
Don't forget about `aes()`!!
</div>

```{r boxplot-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot()
```

```{r boxplot-code-check}
grade_code()
```

### Activity 2: Adding layers

Now, add the individual data point on top of your boxplot

```{r jitter-boxplot, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="jitter-boxplot-hint">
**Dad joke:** Which geom is the most nervous?
</div>

```{r jitter-boxplot-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot() +
  geom_jitter()
```

```{r jitter-boxplot-code-check}
grade_code()
```

### Activity 3: Add colour

Finally, let's add some color to our plot! How would ensure that ONLY the boxplots and NOT the individual data points are colored by according to which state they correspond to?  

Don't worry about the specific colors, just make sure that each state has its own color  

```{r add-color, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="add-color-hint">
**Hint:** Think local, not global
</div>

```{r add-color-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot(aes(colour = state)) +
  geom_jitter()
```

```{r add-color-code-check}
grade_code()
```