---
title: "Themes and Labels Activity"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
gradethis::gradethis_setup()
```
## Themes and Labels Activity

We will continue to use the `midwest` data set for this series of exercises. If you need a scratchpad, feel free to use any of the consoles below.

### Activity 1: Adding labels our boxplot

Again, make a boxplot to plot the distribution of land area for each state, and then add the individual data points  

Then, add labels to the x- and y-axes -  
"State" (x-axis label)  
"Land Area" (y-axis label)  

```{r boxplot-label, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Boxplot"}

```

<div id="boxplot-label-hint">
Use the `labs()` function
</div>

```{r boxplot-label-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot() +
  geom_jitter() +
  labs(x = "State",
       y = "Land Area")
```

```{r boxplot-label-code-check}
grade_code()
```

### Activity 2: Add Color 

Now give all the boxplots and data points different colors depending on which state they correspond to  

```{r boxplot-color, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="boxplot-color-hint">
Use the `color =` parameter
</div>

```{r boxplot-color-solution}
ggplot(data = midwest, mapping = aes(state, area, color = state)) +
  geom_boxplot() +
  geom_jitter()
```

```{r boxplot-color-code-check}
grade_code()
```

### Activity 3: Remove the Legend

Finally, let's remove the color legend. This one is a bit tricky, and you will need to use the `theme()` function. Here's the (extensive!) documentation for this function -  
https://ggplot2.tidyverse.org/reference/theme.html

```{r remove-legend, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="remove-legend-hint">
Use the `legend.position` parameter
</div>

```{r remove-legend-solution}
ggplot(data = midwest, mapping = aes(state, area, color = state)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")
```

```{r remove-legend-code-check}
grade_code()
```