---
title: "ggplot Final Exercise"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
#load libraries
library(learnr)

```

## Optional Exercises

Exercises are opportunities to practices several concepts from a module. These should take 20-30 minutes to complete. The `mtcars`, `mitocarta`, and `scRNAseq` exercises below are roughly equivalent. Choose one and work through it, downloading the answers if needed. Then, try the other example and see if you can complete it without looking at the answer file. To challenge yourself, try the `subcell` example.  
  
To begin, click the download button and then drag the exercise file to your working project directory. Again, the "answer" file shows you one way that you can accomplish the task.   
  
```{r}
h4("mtcars Example")
make_button(file_name = "3-ggplot_exercise_file_mtcars.Rmd", 
            button_style = "primary")
make_button(file_name = "3-ggplot_exercise_file_mtcars_answers.Rmd",
            button_style = "default")
br()
br()
h4("mitocarta Example")
make_button(file_name = "3-ggplot_exercise_file_mitocarta.Rmd", 
            button_style = "primary")
make_button(file_name = "3-ggplot_exercise_file_mitocarta_answers.Rmd",
            button_style = "default")
br()
br()
h4("scRNAseq Example")
make_button(file_name = "3-ggplot_exercise_file_scrnaseq.Rmd",
            button_style = "primary")
make_button(file_name = "3-ggplot_exercise_file_scrnaseq_answers.Rmd",
            button_style = "default")
br()
br()
h4("Subcell Example")
make_button(file_name = "3-ggplot_exercise_file_subcell.Rmd",
            button_style = "primary")
make_button(file_name = "3-ggplot_exercise_file_subcell_answers.Rmd",
            button_style = "default")
actionLink("class3_exercise", "", icon = icon("bug"))

conditionalPanel(
  condition = "input.class3_exercise != 0",
  hr(),
  p("If you are participating in tidybiology live, then this is the file we will use in class. If you are not in a synchronous class, then feel free to use this additional file for practice."), 
  tags$ol(
    tags$li("Create a new project in Rstudio (optional)"), 
    tags$li("Download the project Rmd file (big red button)"), 
    tags$li("Follow along with the exercise")
  ),
  h4("In-class Example"),
  make_button(file_name = "3-ggplot_exercise_file_cmv.Rmd",
              button_style = "danger"), 
  make_button(file_name = "3-ggplot_exercise_file_cmv_answers.Rmd",
              button_style = "default"), 
  hr()
)
```
