---
title: "Exploring Aesthetics using ggplot"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r, include = FALSE}
library(learnr)
gradethis::gradethis_setup()
```
## Exploring Aesthetics Exercise

We will continue to use the `midwest` data set for this series of exercises

### Improving our boxplot

Again, make a boxplot to plot the distribution of land area for each state, and then add the individual data points  

NOTE: From now on, use the "concise" form of ggplot code i.e. Drop `data = ` and `mapping = `

```{r boxplot-redux, exercise = TRUE, exercise.eval = FALSE, exercise.cap = "Boxplot"}

```

<div id="boxplot-redux-hint">
You will need to two geoms for this
</div>

```{r boxplot-redux-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot() +
  geom_jitter()
```

```{r boxplot-redux-code-check}
grade_code()
```

### Adding Color

Make all the individual data points red

```{r boxplot-red, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="boxplot-red-hint">
The additional code goes into one of the two geoms
</div>

```{r boxplot-red-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot() +
  geom_jitter(color = "red")
```

```{r boxplot-red-code-check}
grade_code()
```

### Adjust Transparency

Notice that the points obscure the boxplots. Reduce the opacity of these points by 50%

```{r transparency, exercise = TRUE, exercise.eval = FALSE, exercise.cap = ""}

```

<div id="transparency-hint">
Adjust the alpha parameter
</div>

```{r transparency-solution}
ggplot(data = midwest, mapping = aes(state, area)) +
  geom_boxplot() +
  geom_jitter(color = "red", alpha = 0.5)
```

```{r transparency-code-check}
grade_code()
```