---
title: "ggplot Exercise: mtcars"
output: html_document
---

## Set-up your space
Add a setup chunk that loads the tidyverse packages.

```{r}
```

## Load the mpg dataset
```{r}
```

## First plot
Make a ggplot using the mpg dataset, and compare displacement (x) to miles per gallon (y)
```{r}

```

## Enhance your plot
Add `color`, `size`, `alpha`, or `shape` aesthetics to your graph. Experiment.  
```{r}

```

## Facet
What do `facet_grid()` and `facet_wrap()` do? To figure this out, make a simple plot, such as the one you ran above. Then, you can assign this simple plot to an object, and finally add additional layers (such as your `facets`) to that ggplot object. 
```{r}

```

## Add another variable
Often, data visualizations are rich with information. You previously wrote a ggplot function to only include two variables (x and y). Add a third variable to color
```{r}

```

## Change the geom
Replace this scatterplot with one that draws boxplots. Use the cheatsheet or the internet. Try your best guess.
```{r}
#old

#new

```

## Histograms
Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.
```{r}

```

Use the help page for `geom_histogram` to make the bins 2 units wide.
```{r}

```

Taking the code from the above histogram, make a bar chart for `hwy` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.
```{r}

```

## Final plot
Go back to the beginning to look at the simple plot for the relationship between engine size and mile per gallon. Make a plot that has a scatter plot, and a best fit line using a loess model
```{r}

```

## Make it interactive
Plotly is a package that can take a ggplot and wrap it in a layer of interactivity. Download the plotly package adn then take the last plot you made (scatter plot + fit)  and make an interactive version of it. *Don't forget to load the library after you download it. 

```{r}

```

## Take aways
You can use this code template to make thousands of graphs with **ggplot2**.
```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```