---
title: "ggplot2"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ggplot2: Data visualization

## Install packages
Let's install the following packages: medicaldata, ggplot2, tidyverse, paletter, cowplot, and patchwork needed for this exercise.

Learn more about: [medicaldata](https://higgi13425.github.io/medicaldata/), [ggplot2](https://ggplot2.tidyverse.org/index.html), [paletter](https://github.com/EmilHvitfeldt/paletteer), [cowplot](https://wilkelab.org/cowplot/index.html)

## Load packages
Now, let's load the these packages: medicaldata, ggplot2, tidyverse, paletter, cowplot, and patchwork
```{r}
#install.packages("medicaldata")
library(medicaldata)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("tidyverse)
library(tidyverse)
#install.packages("paletteer")
library(paletteer)
#install.packages("cowplot")
library(cowplot)
#install.packages("patchwork")
library(patchwork)
```

We will be using the `cytomegalovirus` dataset from the medicaldata package. Load the dataset and assign it to the following object, `cyto.` Hint: Use the `::` to select for the dataset.
```{r}
cyto <- medicaldata::cytomegalovirus
```

Let's take a glimpse of the `cyto` dataset.
```{r}
glimpse(cyto)
```

## ggplot2
In this section we will be using ggplot to create data visualizations.

## ggplot: geom_bar
Let's create bar plot with our `cyto` dataset to determine the number of patients per `diagnosis`. Hint: Do not forget to include `geom` for the aesthetics. 
```{r}
 ggplot(data = cyto) + # call our data
  geom_bar(mapping = aes(x = diagnosis)) # add elements
```

With the previous script, let's `fill` to the bar plot with color by `diagnosis`. Hint: edit the aesthetic of the bar plot 
```{r}
ggplot(data = cyto) + 
  geom_bar(mapping = aes(x = diagnosis, fill = diagnosis))
```

## Change color scheme:
Let's add to the script above and use the `miscpalettes::brightPastel` palette from the `paletteer` package to change the color scheme of our graph. Hint: Include the `scale_fill_paletteer_d()` function to make this change. 
```{r}
ggplot(data = cyto) +
  geom_bar(mapping = aes(x = diagnosis, fill = diagnosis)) +
  scale_fill_paletteer_d("miscpalettes::brightPastel") # from the paletteer package
```

## Change layout - Theme:
Now we will be focusing on changing the theme of our bar plot! Continuing from the same script above, let's:
1. Change the `theme` of the bar plot to black and white (bw)
2. Edit the `theme`, with the `theme()` function to:
  2a. Tilt the text on the x-axis to an `angle` of `60` degrees and horizontal adjustment (`hjust`) to `1`.         Hint: Use `axis.text.x = element_text()` from the `theme()` function
  2b. Remove the title of the x-axis. Hint: Incorporate `axis.title.x = element_blank()` from the `theme()`         function
  2c. To remove the legend, change the `legend.position` to `none`
3. Change the title of the y-axis to `Number of patients`. Hint: Use the `labs()` function
  
```{r}
ggplot(data = cyto) + # call our data
  geom_bar(mapping = aes(x = diagnosis, fill = diagnosis)) +
  scale_fill_paletteer_d("miscpalettes::brightPastel") + # changes the color scheme
  theme_bw() + # changes the theme to black and white (bw)
  theme(axis.text.x = element_text(angle = 60, hjust = 1), # tilts the text of the x-axis
        axis.title.x = element_blank(), # removes the title of the x-axis
        legend.position = "none") + # removes the legend
  labs(y = "Number of Patients") # changing the label for y-axis
```

### ggplot: facet_grid
Say we want to see if there are any differences in number of patient diagnoses based on sex, we will add the `facet_grid()` function to make this observation. However, since the `sex` variable in our `cyto` dataset is coded by numbers, 0 = Female and 1 = Male, we will have to store this information in an object called `plot_names`. The `plot_names` object will be used as our new `labeller` for the `facet_grid()` function.
```{r}
plot_names <- c("0" = "Female", # to create new title names for panels
                "1" = "Male")

ggplot(data = cyto) +
  geom_bar(mapping = aes(x = diagnosis, fill = diagnosis)) +
  scale_fill_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none") +
  labs(y = "Number of Patients") +
  facet_grid(cols = vars(sex), labeller = as_labeller(plot_names)) # use labeller to change the title of the panels
```

### ggplot: geom_jitter
Now we will use another `geom` aesthetic called `jitter` in which we will compare the `CD34.dose` to `CD3.dose` and color the jitters based by `diagnosis`. 
Let's also change the:
1. Color palette to `miscpalettes::brightPastel` 
2. Theme to black and white (bw)
3. Text on the x-axis to contain an `angle` of `60` degrees and horizontal adjustment (`hjust`) to `1`.
4. Title of the x-axis to `CD34 Dose`and y-axis to `CD3 Dose`

```{r}
ggplot(data = cyto) +
  geom_jitter(mapping = aes(x = CD34.dose, y = CD3.dose, color = diagnosis)) +
  scale_color_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none") +
  labs(x = "CD34 Dose",
       y = "CD3 Dose")
```

### ggplot: geom_boxplot
Next, with the `cyto` dataset, we will be creating a simple box plot looking the `diagnosis` to `TNC.dose`. 
```{r}
ggplot(data = cyto) +
  geom_boxplot(mapping = aes(x = diagnosis, y = TNC.dose))
```

### ggplot: combining a box and jitter plot
Using the `cyto` dataset, let's create a box plot looking the `diagnosis` to `TNC.dose`, colored by `diagnosis`. Next, add a jitter plot with the same conditions.
Let's also change the:
1. Color palette of the plot to `miscpalettes::brightPastel` 
2. Theme to black and white (bw)
3. Text on the x-axis to contain an `angle` of `60` degrees and horizontal adjustment (`hjust`) to `1`.Remove     the title of the x-axis and legend.
4. Title of the y-axis to `TNC Dose`
```{r}
ggplot(data = cyto) + 
  geom_boxplot(mapping = aes(x = diagnosis, y = TNC.dose,
                             color = diagnosis)) +
  geom_jitter(mapping = aes(x = diagnosis, y = TNC.dose,
                            color = diagnosis), alpha = 0.5) + # alpha lightens the jitter color
  scale_fill_paletteer_d("miscpalettes::brightPastel") +
  scale_color_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none") +
  labs(y = "TNC Dose")
```

### Another package to check out: Patchwork
Patchwork is a great package used to combine multiple plots into one! 
Preparation for patchwork:
Store three of the ggplots into three objects: `NoP` will contain the 1st chunk of code from the Change layout - Theme section, `CD` will be the chunk of code from the ggplot: geom_jitter section, the code from ggplot: combining a box and jitter plot, will be stored in `TNC_P`.

```{r}
NoP <- ggplot(data = cyto) +
  geom_bar(mapping = aes(x = diagnosis, fill = diagnosis)) +
  scale_fill_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none") +
  labs(y = "Number of Patients")

CD <- ggplot(data = cyto) +
  geom_jitter(mapping = aes(x = CD34.dose, y = CD3.dose, color = diagnosis)) +
  scale_color_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  labs(x = "CD34 Dose",
       y = "CD3 Dose")

TNC_P <- ggplot(data = cyto) + 
  geom_boxplot(mapping = aes(x = diagnosis, y = TNC.dose,
                             color = diagnosis)) +
  geom_jitter(mapping = aes(x = diagnosis, y = TNC.dose,
                            color = diagnosis), alpha = 0.5) +
  scale_fill_paletteer_d("miscpalettes::brightPastel") +
  scale_color_paletteer_d("miscpalettes::brightPastel") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none") +
  labs(y = "TNC Dose")
```

### Patchwork
Run the code below to see for yourself:
```{r}
(NoP | TNC_P) / CD + # objects we created from our previous chunk of code
  plot_annotation(tag_levels = c("A", "B", "C")) & # adding labels to our plots
  theme(plot.tag = element_text(size = 12)) # change the size of the text
```

