---
title: "Ggplot2 Exercises"
output: html_document
author: Junqi Lu
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
# install.packages("ggExtra")
library(tidyverse)
library(ggExtra)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `scRNAseq`

```{r}
#install.packages("remotes")
#remotes::install_github("devangthakkar/scRNAseq", force = TRUE)
library(scRNAseq)
data(scRNAseq)
```

Take a quick look at the contents of this dataset before attempting the questions below

```{r}

```

*WARNING*: These data are MADE UP. Please therefore only use this dataset for this exercise!

## Please answer the following questions

What is the distribution of transcript lengths for the genes in this dataset? Hint - the relevant variable is `transcript_length`

```{r}

```

How many *unique* organs are represented in this dataset?

```{r}

```

Create a new data frame that only contains genes whose transcript length is less that 7500 bp long. Also discard all rows that contain "other" in the `organ` column.

Call this new data frame `three_organ_df`

```{r}

```

Create boxplots for transcript lengths for each organ type. Overlay the individual points on top of these boxplots

```{r}

```

Give the previous plot appropriate labels for the x- and y-axes, and a title

```{r}

```

Many of the points are obscuring each other. What's one way to "fix" this?

```{r}

```

How would you make all the black points red?

```{r}

```

Plot the gene expression values for `cell_1` against `cell_2`. Make the points blue, change their size to 1.5, and adjust the transparency to 0.3

Also, add a diagonal reference line (45 degrees) that intersects with the intercept. Set the colour of this line to green

```{r}

```

Change the plot above so that data points corresponding to different organ types have different colours

```{r}

```

Split the plot above into three plots that correspond to the three organ types. Bonus: Have the label of each plot be in this format - organ: organ_name e.g. "organ: brain"

```{r}

```

Using the un-split version of the previous plot, add marginal histograms above and to the right of this plot. Hint - use the `ggMarginal()` function from the `ggExtra()` package

```{r}

```

Give the previous plot appropriate labels for the x- and y-axes, and a title. Also make the x- and y-axis text bold, and remove the colour legend

```{r}

```
